Determinism, Complexity, and Predictability in Computer Performance

### 2014-01-11 pecomplexity.tex
clean up text some

The weighted probability of a permutation is then:
\begin{align*}
  p_w(\pi) = \frac{\displaystyle \sum_{t \le T - n} w(x_{t+1:t+n}) \cdot \delta(\phi(x_{t:t+n}), \pi) }{\displaystyle \sum_{t \le T - n} w(x_{t+1:t+n})}
\end{align*}
where $\delta(x, y)$ is 1 if $x = y$ and 0 otherwise. Effectively, this weighted
probability enhances permutations involved in ``large'' features and demotes
permutations which are small in amplitude relative to the features of the time
series. The weighted permutation entropy is then:
\begin{align*}
  H_w(n) = - \sum_{\pi \in \mathcal{S}_n} p_w(\pi) \log_2 p_w(\pi)
\end{align*}

@@

The relationship between prediction accuracy and the permutation
entropy (PE) is as we conjectured: performance traces with high
PE---those whose temporal complexity is high, in the sense that little
information is being propagated forward in time---are indeed harder to
predict using the simple deterministic forecast model described in the
previous section.  The effects of changing $n$ are also interesting:
using a longer wordlength generally lowers the PE---a natural
consequence of finite-length data---but the falloff is less rapid in
some traces than in others, suggesting that those values are closer to
the theoretical asymptote that exists for perfect data.  The
persistent PE values of 0.5--0.6 for the {\tt row\_major} and {\tt
  col\_major} cache-miss traces are consistent with dynamical chaos,
further corroborating the results of~\cite{mytkowicz09}.  (PE values
above 0.97 are consistent with white noise.)  Interestingly, the
processor-load traces for these two microkernels exhibit more temporal
complexity than the cache-miss traces.  This may be a consequence of
the lower baseline value of this time series.

>>>>

The weighted probability of a permutation is then:
\begin{align*}
  p_w(\pi) = \frac{\displaystyle \sum_{t \le T - n} w(x_{t+1:t+n}) \cdot \delta(\phi(x_{t:t+n}), \pi) }{\displaystyle \sum_{t \le T - n} w(x_{t+1:t+n})}
\end{align*}
where $\delta(x, y)$ is 1 if $x = y$ and 0 otherwise. Effectively, this weighted
probability enhances permutations involved in ``large'' features and demotes
permutations which are small in amplitude relative to the features of the time
series. The weighted permutation entropy is then:
\begin{align*}
  H_w(n) = - \sum_{\pi \in \mathcal{S}_n} p_w(\pi) \log_2 p_w(\pi),
\end{align*}
which can also be normalized by dividing by $\log_2(n!)$, and will be in all the
results of this paper.

@@

The relationship between prediction accuracy and the weighted permutation
entropy (WPE) is much as we conjectured: performance traces with high WPE ---
those whose temporal complexity is high, in the sense that the entropy
generation per value is approximately equal to the total entropy per value thus
little information is being propagated forward in time --- are indeed harder to
predict using the simple deterministic forecast model described in the previous
section.  The effects of changing the word length $n$ are also interesting:
using a longer word length generally lowers the WPE --- a natural consequence of
finite-length data --- but the falloff is less rapid in some traces than in
others, suggesting that those values are closer to the theoretical asymptote
that exists for perfect data. The persistent WPE values of 0.5--0.6 for the {\tt
col\_major} trace are consistent with dynamical chaos, further corroborating the
results of~\cite{mytkowicz09}.
